`infercnv` results exploration for `r params$sample_id`

### 2024-11-15 analyses/cell-type-wilms-tumor-06/notebook_template/06_cnv_infercnv_exploration.Rmd
fix typos on 06 notebook, but do not regenerate

`infercnv` currently support two models for HMM-based CNV prediction, what we refer to as the i3 and i6 models. These are set in the 'infercnv::run()' as HMM_type='i3' or HMM_type='i6' (i6 is default). Each method operates on the 'preliminary infercnv object' which has been processed through the standard inferCNV processing routines, involving subtraction of signal corresponding to 'normal (reference)' cells and smoothing operations.

@@

For a `Seurat` object an `infercnv` object created with the script `06_infercnv.R` using `reference_value` as a reference, the function `Do_CNV_score` calculate a CNV score per cell.
The score is calculated based on the [biostar discussion](https://www.biostars.org/p/9573777/).
The function `Do_CNV_score` returns the `Seurat` object with an additional metadata named `CNV-score_{reference_value}.`

@@

#### Visualize seurat clusters and metadata

For a Seurat object `object`and a metadata `metadata`, the function `visualize_metadata` will plot `FeaturePlot` and `BarPlot`

@@

#### Wrapper function to explore `infercnv` HMM cnv prediction

The `wrapper_explore_hmm` take as input the `infercnv_obj` generated with `infercnv` HMM cnv predictions.
The wrapper allows the following steps and plots:

@@

#####  Distribution of CNV estimation in the Wilms tumor copartments

@@

##### DotPlot

The `Dotplot` representation summarizes the percentage of cells in each compartment with cnv in each of the 22 chromosomes.

@@

### Heatmap of infercnv results

Here we plot the output of `infercnv` as heatmaps of CNV.
We first look at the png file generated by the `infercnv` function.
We then used the `infercnv object` to look at mean CNV value across compartments (immune, endothelial, stroma and fetal nephron).

@@

```{r fig.width=20, fig.height=5, out.width='100%', out.height='100%'}
infercnv_obj <- list()
for (reference_value in c("reference-none", "reference-immune", "reference-endothelium", "reference-both")) {
  infercnv_obj[[reference_value]] <- readRDS(file.path(result_dir, glue::glue(reference_value, "_HMM-no"), glue::glue("06_infercnv_", params$sample_id, "_", reference_value, "_HMM-no.rds")))
  print(Do_CNV_heatmap(object = srat, infercnv_obj = infercnv_obj[[reference_value]], group.by = "fetal_kidney_predicted.compartment", reference_value = reference_value))
}
```
These heatmaps emphasize the importance of the selection of normal cells prior the inference of CNV.
The normal reference should contain as much cell types as possible, in order to minimize false positive CNV.
In our case, we should take immune and entodethial cells when possible.

@@

We want to calculate a single CNV score and asess if/how it can be use to define cells with CNV versus stable/normal cells. 
We defined the score as discribed in the [biostar discussion](https://www.biostars.org/p/9573777/).

@@

We then explore infercnv results generated with immune and endothelial cells as reference, using a [HMM-i3 prediction models](https://github.com/broadinstitute/infercnv/wiki/infercnv-i3-HMM-type).

@@

and explore the CNV results using the `wrapper_explore_hmm` fucntion.

@@

## Comparisons of inter- and intra-patient global cnv score with HMM prediction model

@@

- HMM-i3 with inter-patient endothelial and immunce cells as reference

- HMM-i3 with intra-patient endothelial and immunce cells as reference

- HMM-i6 with intra-patient endothelial and immunce cells as reference

@@

   + HMM-i3 with inter-patient endothelial and immunce cells as reference

   + HMM-i3 with intra-patient endothelial and immunce cells as reference

   + HMM-i6 with intra-patient endothelial and immunce cells as reference

@@

   + The HMM-i3 model with inter-patient endothelial and immunce cells as reference has the advantage to be usable for all Wilms tumor samples, including the ones with a very low number of immune and/or endothelial cells.

   + The HMM-i3 model with intra-patient endothelial and immunce cells as reference seems to be the cleaner, ~fast to run (10 minutes per samples) and is more precise than the HMM-i3 with the inter-patient reference.

   + The HMM-i6 model with intra-patient endothelial and immunce cells as reference is very slow (~2 hours per sample) and couldn't be used for the entire cohorte.
It is more noisy than the i3 version. However, it could have the potential to detect cancer cells with very low CNV profile.

- Surprisingly, running `infercnv` with emdothelial and immune cells from (i) the same patient or (ii) a set of Wilms tumor patients do not seem to affect drastically the results. 
Some false positive CNV might occur in every patient due to the inter-patient variability. 
By comparing the results in conditions (i) and (ii), we should be able to understand which false positive are recurrent and do not take them into account.

>>>>

`infercnv` currently support two models for HMM-based CNV prediction, what we refer to as the i3 and i6 models. These are set in the `infercnv::run()` as HMM_type='i3' or HMM_type='i6' (i6 is default). Each method operates on the preliminary `infercnv` object which has been processed through the standard inferCNV processing routines, involving subtraction of signal corresponding to "normal (reference)" cells and smoothing operations.

@@

For a `Seurat` object an `infercnv` object created with the script `06_infercnv.R` using `reference_value` as a reference, the function `Do_CNV_score` calculate a CNV score per cell.
The score is calculated based on the [this discussion](https://www.biostars.org/p/9573777/).
The function `Do_CNV_score` returns the `Seurat` object with an additional metadata named `CNV-score_{reference_value}.`

@@

#### Visualize Seurat clusters and metadata

For a Seurat object `object` and a metadata `metadata`, the function `visualize_metadata` will plot `FeaturePlot` and `BarPlot`

@@

#### Wrapper function to explore `infercnv` HMM CNV prediction

The `wrapper_explore_hmm` take as input the `infercnv_obj` generated with `infercnv` HMM CNV predictions.
The wrapper allows the following steps and plots:

@@

#####  Distribution of CNV estimation in the Wilms tumor compartments

@@

##### Dot Plot

The `Dotplot` representation summarizes the percentage of cells in each compartment with CNV in each of the 22 chromosomes.

@@

### Heatmap of `infercnv` results

Here we plot the output of `infercnv` as heatmaps of CNV.
We first look at the PNG file generated by the `infercnv` function.
We then used the `infercnv object` to look at mean CNV value across compartments (immune, endothelial, stroma and fetal nephron).

@@

```{r fig.width=20, fig.height=5, out.width='100%', out.height='100%'}
infercnv_obj <- list()
for (reference_value in c("reference-none", "reference-immune", "reference-endothelium", "reference-both")) {
  infercnv_obj[[reference_value]] <- readRDS(file.path(result_dir, glue::glue(reference_value, "_HMM-no"), glue::glue("06_infercnv_", params$sample_id, "_", reference_value, "_HMM-no.rds")))
  print(Do_CNV_heatmap(object = srat, infercnv_obj = infercnv_obj[[reference_value]], group.by = "fetal_kidney_predicted.compartment", reference_value = reference_value))
}
```
These heatmaps emphasize the importance of the selection of normal cells prior the inference of CNV.
The normal reference should contain as much cell types as possible, in order to minimize false positive CNV.
In our case, we should take immune and endothelial cells when possible.

@@

We want to calculate a single CNV score and asesss if/how it can be use to define cells with CNV versus stable/normal cells. 
We defined the score as described in the [biostar discussion](https://www.biostars.org/p/9573777/).

@@

We then explore `infercnv` results generated with immune and endothelial cells as reference, using a [HMM-i3 prediction models](https://github.com/broadinstitute/infercnv/wiki/infercnv-i3-HMM-type).

@@

and explore the CNV results using the `wrapper_explore_hmm` function.

@@

## Comparisons of inter- and intra-patient global CNV score with HMM prediction model

@@

- HMM-i3 with inter-patient endothelial and immune cells as reference

- HMM-i3 with intra-patient endothelial and immune cells as reference

- HMM-i6 with intra-patient endothelial and immune cells as reference

@@

   + HMM-i3 with inter-patient endothelial and immune cells as reference

   + HMM-i3 with intra-patient endothelial and immune cells as reference

   + HMM-i6 with intra-patient endothelial and immune cells as reference

@@

   + The HMM-i3 model with inter-patient endothelial and immune cells as reference has the advantage to be usable for all Wilms tumor samples, including the ones with a very low number of immune and/or endothelial cells.

   + The HMM-i3 model with intra-patient endothelial and immune cells as reference seems to be the cleaner, ~fast to run (10 minutes per samples) and is more precise than the HMM-i3 with the inter-patient reference.

   + The HMM-i6 model with intra-patient endothelial and immune cells as reference is very slow (~2 hours per sample) and couldn't be used for the entire cohort.
It is more noisy than the i3 version. However, it could have the potential to detect cancer cells with very low CNV profile.

- Surprisingly, running `infercnv` with endothelial and immune cells from (i) the same patient or (ii) a set of Wilms tumor patients do not seem to affect drastically the results. 
Some false positive CNV might occur in every patient due to the inter-patient variability. 
By comparing the results in conditions (i) and (ii), we should be able to understand which false positive are recurrent and do not take them into account.
